---
title: "Comparaison d'isolats bactûˋriens par sûˋquenûÏage gûˋnomique complet"
subtitle: ""
author: "Equipe PHE3ID"
date: "`r format(Sys.time(), '%d %B %Y')`"
params:
  cluster_id: ""
  species: ""
  replicon: "Chromosome"
  env_samples: ""
  path_to_reports: "2024-01-25_clone_reports.Rdata"
  path_to_scan:  "2024-01-25_annotation_scan.Rdata"
output:
  pdf_document:
      toc: true
      toc_depth: 2
      number_sections: true
fontsize: 12pt
#classoption: landscape
header-includes:
- \usepackage{graphicx}
- \usepackage{fancyhdr}
- \usepackage{float}
- \usepackage{titling}
- \usepackage{svg}
- \usepackage{booktabs}
- \usepackage[table]{xcolor}
- \pagestyle{fancy}
- \fancyhead[L]{Equipe PHE3ID}
- \fancyhead[C]{Cluster `r params$cluster_id` - `r params$replicon` - `r params$species`}
- \fancyhead[R]{\includegraphics[width=30pt]{pictures/logo_hcl.png}}
- \setlength{\headheight}{35pt}
- \addtolength{\topmargin}{-19pt}
- \pretitle{\begin{center}
  \hfill
  \includegraphics[width=1in,height=1in]{pictures/logo_hcl.png}\LARGE\\}
- \posttitle{\end{center}}
---

<!--
*******************************************************************
Title: Non-interactive_quick_cluster_report.Rmd
Description: Report at species level, comparing isolates with SNP calling and MLST annotation.
Authors: Aurûˋlie Fischer
Research team: PHE3ID, CIRI, Hospices Civils de Lyon
License: GNU Affero General Public License v3.0 (AGPL-3.0)
SPDX-License-Identifier: AGPL-3.0-only
Copyright (C) 2026 Aurûˋlie Fischer
*******************************************************************
-->

\newpage

Ce rapport d'ûˋtudes de cluster ûˋtudie les isolats de l'espû´ce ___ XXX ___ du service YYYY.

```{r setup, include=FALSE}
# Some cleaning [no: if cleaning, then we cannot use params data]
# rm(list = ls())

# Setting R chunks options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.align = 'center')

# Loading libraries
usePackage <- function(p) {
  if (!is.element(p, installed.packages()[,1]))
    install.packages(p, dep = TRUE)
  require(p, character.only = TRUE)
}
libraries <- c("lubridate", "data.table", "stringr", "knitr", "ggplot2", "dplyr", "scales", "tidyr", "kableExtra", "ggpubr", "RColorBrewer", "tidyverse")
for (lib in libraries){usePackage(lib)}


# Loading Data, some clusters are to be compared to others
load(params$path_to_reports)
if(params$cluster_id==2){
  reports <- final_data[cluster_id %in% c(2, 161)]
} else if(params$cluster_id==348){
  reports <- final_data[cluster_id %in% c(348, 359)]
} else if(params$cluster_id==316){
  reports <- final_data[cluster_id %in% c(316, 322)]
} else if(params$cluster_id==413){
  reports <- final_data[cluster_id %in% c(413, "406-413")]
} else if(params$cluster_id==393){
  reports <- final_data[cluster_id %in% c(393, 392, 245)]
} else {
  reports <- final_data[cluster_id == params$cluster_id]
}

# Studied species
if(params$species=="chromosome"){
  species <- "-"
} else {
  species <- params$species
}
# In case of ecloacae, include cronobacter !
if(params$species=="ecloacae"){
  if(params$cluster_id==2){
    # Take also E. aerogenes
    reports <- reports[mlst_species == species | mlst_species == "kaerogenes",]
  } else {
    # Also include cronobacter
    reports <- reports[mlst_species == species | mlst_species == "cronobacter",]
  }
} else if(params$species=="ecoli"){
  # Also include ecoli_achtman_4
  reports <- reports[mlst_species == species | mlst_species == "ecoli_achtman_4" | mlst_species == "escherichia",]
} else if(params$species=="kpneumoniae"){
  # Also include klebsiella
  reports <- reports[mlst_species == species | mlst_species == "klebsiella",]
} else if(params$cluster_id %in% c(420, 480)){
  # Take all three samples
  reports <- reports
} else {
  reports <- reports[mlst_species == species,]
}
rm(final_data)
reports <- reports %>%
  filter(genome!="Epi-205" | genome != "Epi-615" | genome != "Epi-2348")
# Global ggplot theme settings
global_theme <- theme(plot.title=element_text(size=16, face="plain", hjust = 0.5, vjust=2),
            axis.text.x=element_text(size=9),
            axis.text.y=element_text(size=10),
            axis.title.x=element_text(size=12, vjust = -1),
            axis.title.y=element_text(size=12, angle = 90, vjust = 2),
            legend.title=element_text(size=12),
            legend.text = element_text(size=9))
theme_set(global_theme)

rm(lib, libraries, usePackage)
```

# Description des prûˋlû´vements

```{r global_infos}
# Calcul de la date de prûˋlû´vement la + prûˋcoce
earliest <- format(min(reports$sampling_date), "%d/%m/%Y")
# Calcul de la date de prûˋlû´vement la + tardive
latest <- format(max(reports$sampling_date), "%d/%m/%Y")
# Nombre d'ûˋchantillons dans le cluster
nb_samples <- length(reports$sampling_date)
# Nombre de patients concernûˋs
nb_patients <- length(unique(reports$ipp))
# Nombre d'ûˋchantillons environnement s'il y en a
nb_env_samples <- 0
if(!(params$env_samples=="")){
  nb_env_samples <- length(reports[reports$ipp == params$env_samples]$genome)
  nb_patients <- nb_patients - 1
}

```

```{r general_desc, results='asis'}
# Global description of cluster isolates, patients/environment, counts, dates
if(nb_patients==1){
  nb_patients <- paste(nb_patients, " patient")
} else {
  nb_patients <- paste(nb_patients, "patients")
}

if(nb_env_samples != 0){
  if(earliest==latest){
    cat("Groupe de ",nb_samples," isolats suspects chez ",nb_patients," et ",nb_env_samples," prûˋlû´vements d'environnement, prûˋlevûˋs le ", earliest, ".")
  } else {
    cat("Groupe de ",nb_samples," isolats suspects chez ",nb_patients," et ",nb_env_samples," prûˋlû´vements d'environnement, prûˋlevûˋs du ", earliest," au ",latest,".")
  }
} else {
  if(earliest==latest){
      cat("Groupe de ",nb_samples," isolats suspects chez ",nb_patients,", prûˋlevûˋ(s) le ", earliest, ".")
  } else {
      cat("Groupe de ",nb_samples," isolats suspects chez ",nb_patients,", prûˋlevûˋ(s) du ", earliest," au ",latest,".")
  }


}
rm(nb_env_samples, nb_samples, nb_patients, earliest, latest)
```

```{r dates, out.height="300pt"}
# Computing limits of histogram (earliest and latest months)
first_month <- floor_date(min(reports$sampling_date), "month") %m-% months(1)
last_month <- ceiling_date(max(reports$sampling_date), "month")
months_period <- interval(ymd(first_month),ymd(last_month)) %/% months(1)
if(months_period < 6){
  first_month <- as.Date("2025-06-01")
  last_month <- as.Date("2026-06-01")
}
# Species in title
{
if(params$cluster_id %in% c(3, 395)){
  species_in_title <- "S. marcescens"
} else if(params$cluster_id %in% c(96, 113, 121, 127, 189, 226, 240, 260, 286, 295, 324, 334, 374, 361, 485, 524, 572, 607, 591, 630)){
  species_in_title <- "S. capitis"
} else if(params$species == "cfreundii" & params$cluster_id == 39) {
  species_in_title <- "C. freundii/braakii/youngae"
} else if(params$species == "escherichia") {
  species_in_title <- "E. coli"
} else if(params$species == "ecloacae" & params$cluster_id == 39) {
  species_in_title <- "E. cloacae/hormaechei"
} else if(params$cluster_id == 50){
  species_in_title <- "A. faecalis"
} else if(params$cluster_id == 48){
  species_in_title <- "B. fragilis"
} else if(params$cluster_id %in% c(60, 115)){
  species_in_title <- "A. ursingii"
} else if(params$cluster_id %in% c(40, 228, 414)){
  species_in_title <- "P. mirabilis"
} else if(params$cluster_id %in% c(99, 122, 173, 335, 531)){
  species_in_title <- "A. xylosidans"
} else if(params$cluster_id %in% c(87, 310)){
  species_in_title <- "A. nosocomialis"
} else if(params$cluster_id %in% c(46, 315, 470)){
  species_in_title <- "A. baumannii"
} else if(params$cluster_id %in% c(73, 508)){
  species_in_title <- "S. saprophyticus"
} else if(params$cluster_id == 75){
  species_in_title <- "S. sonnei"
} else if(params$cluster_id == 111){
  species_in_title <- "M. catharralis"
} else if(params$cluster_id %in% c(110, 119)){
  species_in_title <- "S. constellatus"
} else if(params$cluster_id %in% c(138, 190)){
  species_in_title <- "T. otitidis"
} else if(params$cluster_id %in% c(141, 375, 501, 608, 634)){
  species_in_title <- "H. alvei"
} else if(params$cluster_id == 147){
  species_in_title <- "C. braakii"
} else if(params$cluster_id == 136){
  species_in_title <- "L. rhamnosus"
} else if(params$species == "klebsiella") {
  species_in_title <- "K. pneumoniae"
} else if(params$cluster_id %in% c("165", "234", "594")) {
  species_in_title <- "E. asburiae"
} else if(params$cluster_id == "172") {
  species_in_title <- "D. acidovorans"
} else if(params$cluster_id %in% c("174","267", "436")) {
  species_in_title <- "S. caprae"
} else if(params$cluster_id == "181") {
  species_in_title <- "E. cloacae/E. asburiae"
} else if(params$cluster_id == "285") {
  species_in_title <- "A. johnsonii"
} else if(params$cluster_id == "292") {
  species_in_title <- "P. mirabilis"
} else if(params$cluster_id == "287") {
  species_in_title <- "S. simulans"
} else if(params$cluster_id %in% c("288", "482")) {
  species_in_title <- "S. anginosus"
} else if(params$cluster_id == "301") {
  species_in_title <- "M. morganii"
} else if(params$cluster_id == "306") {
  species_in_title <- "S. constellatus"
} else if(params$species == "staphlugdunensis") {
  species_in_title <- "S. lugdunensis"
} else if(params$cluster_id %in% c("349", "562")) {
  species_in_title <- "S. pettenkoferi"
} else if(params$cluster_id == "360") {
  species_in_title <- "S. lactis/paracarnis"
} else if(params$cluster_id == "371") {
  species_in_title <- "R. radiobacter"
} else if(params$cluster_id %in% c("355", "559")) {
  species_in_title <- "C. striatum"
} else if(params$cluster_id == "343") {
  species_in_title <- "P. citronellolis"
} else if(params$cluster_id %in% c("429", "532")) {
  species_in_title <- "P. acnes"
} else if(params$cluster_id == "406") {
  species_in_title <- "K. pneumoniae"
} else if(params$cluster_id == "440") {
  species_in_title <- "B. thuringiensis"
} else if(params$cluster_id == "446") {
  species_in_title <- "P. putida"
} else if(params$species == "neisseria") {
  species_in_title <- "N. gonorrhoeae"
} else if(params$cluster_id == "498") {
  species_in_title <- "P. penneri"
} else if(params$cluster_id == "584") {
  species_in_title <- "S. pasteuri"
} else if(params$cluster_id == "578") {
  species_in_title <- "C. youngae"
} else if(params$cluster_id == "628") {
  species_in_title <- "P. rettgeri"
} else {
  species_in_title <- gsub("^([A-Z])", "\\1. ", str_to_title(params$species))
}
}

# If there are some environmental samples in the group
# Break data by month, e.g one bin one month
out <- reports %>%
  select(genome, sampling_date, ipp) %>%
  mutate(month = floor_date(sampling_date, "month")) %>%
  mutate(patients=ifelse(ipp==params$env_samples, "environnement", "patient")) %>%
  select(-ipp)
ggplot_title <- bquote(atop("Histogramme des dates de prûˋlû´vement", 
                            "de l'espû´ce"~italic(.(species_in_title))))
# Plotting as a barplot of month categories
g <- ggplot(data=out, aes(x=month, fill = patients)) +
   geom_bar(stat="count", aes(text = paste('Nombre: ', after_stat(count))), width = 20) +
   ggtitle(ggplot_title) +
   labs(x = "Date de prûˋlû´vement par mois", y = "Nombre d'ûˋchantillons", fill = "Type de prûˋlû´vement") +
  scale_x_date(date_breaks = "month", date_labels = "%m/%Y", date_minor_breaks = "month", 
               limits = c(first_month, last_month)) +
  scale_y_continuous(n.breaks = 10) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1, size = 8))

g

rm(first_month, last_month, out, g, species_in_title, months_period)
```
\newpage

# Etudes comparatives basûˋes sur les SNPs

```{r which_dist}
# Working on second scale of SNPs[10 or 15 SNPs generally]
clust_dist <- sort(str_match(colnames(reports), "clust_dist_\\d+"))[2]
nb_snps <- str_extract(clust_dist, "\\d+")
```

Deux isolats sont clonaux s'ils sont distants de moins **de `r nb_snps` SNPs**. Les isolats clonaux sont regroupûˋs en cluster.

## Bilan par clone identifiûˋ

```{r clone, out.height="300pt"}
# Get data and arrange ipp/text
out <- reports %>%
  mutate(ipp=ifelse(ipp==params$env_samples, "environnement", ipp)) %>%
  mutate(text_on_bar=ifelse(ipp=="environnement", "environnement", ""))

# Cluster refactoring 
tmp <- unique(out[[clust_dist]])
clust_levels <- tmp[order(nchar(tmp), tmp)]
out[[clust_dist]] <- factor(out[[clust_dist]], levels = clust_levels)
# Color editing
cbPalette <- colorRampPalette(brewer.pal(9, "Set1"))

# Barplot of number of samples per clone
g <- ggplot(data=out, aes(x=as.factor(eval(as.name(clust_dist))), fill = ipp)) +
      geom_bar(stat="count")+
      stat_count(geom = "text", colour = "white", size = 3, angle = 0,
                 aes(label = ipp),position=position_stack(vjust=0.5)) +
      ggtitle(str_replace(clust_dist, "clust_dist_", "Clustering Distance ")) +
      labs(x = "Clusters", y = "Nombre d'ûˋchantillons", fill = "Identifiant Patient (IPP)") +
      scale_y_continuous(n.breaks = 5)+
      scale_fill_manual(values=cbPalette(length(unique(reports$ipp))))+
      guides(fill=guide_legend(ncol=2)) +
      theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1, size = 6))

g

rm(g, cbPalette, tmp, tmp2)
```

```{r desc_clust, results='asis'}
# Description of identified clusters (counts + patient/environment)
clusters <- names(which(table(reports[[clust_dist]]) > 1))
samples_clustered <- out[which(out[[clust_dist]] %in% clusters)]
nb_clusters <- length(clusters)
if(nb_clusters==1){
  cat("Au seuil de ", nb_snps," SNPs, 1 cluster d'isolats est formûˋ.\n\n")
} else if(nb_clusters==0){
  cat("Au seuil de ", nb_snps," SNPs, aucun cluster d'isolats n'est formûˋ.\n\n")
} else {
  cat("Au seuil de ", nb_snps," SNPs, ", nb_clusters, " clusters d'isolats sont formûˋs.\n\n")
}
for(cl in clusters){
  cl_samples <- out[which(out[[clust_dist]] == cl)]
  cat("* Le cluster ", cl, " : groupe de ", length(cl_samples$genome), " isolats prûˋlevûˋs ")
  if("environnement" %in% cl_samples$ipp){
    # If there are environmental samples
    nb_env <- length(cl_samples[cl_samples$ipp == "environnement"]$genome)
    cat("de l'environnement")
    if(length(grep("\\d+", cl_samples$ipp)) > 0) {cat(" et ")}
  }
  if(length(grep("\\d+", cl_samples$ipp)) > 0){
    # If there are samples from patients
    nb_patients <- length(unique(cl_samples[cl_samples$ipp != "environnement"]$ipp))
    if(nb_patients > 1){
      cat("de ", nb_patients, " patients")
    } else {
      cat("d'un patient")
    }
  }
  cat(".\n\n")
}

rm(out, samples_clustered, cl, clusters, nb_clusters)
```
\newpage

## Bilan par prûˋlû´vement

La description des isolats ûˋtudiûˋs :

```{r final_table}
# Final recap table including results of clustering for all isolates
# Arranging data
tab <- reports %>%
  select(glims, ipp, sampling_date, type_prlvt, soc_species, sequence_type,
         as.name(clust_dist)) %>%
  arrange(nchar(eval(as.name(clust_dist))), eval(as.name(clust_dist))) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="SURV", "Surveillance de flore", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="ECBU", "Examen cytobactûˋriologique des urines", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="BPO", "Broncho-pulmonaire", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="CATHE", "Cathûˋter", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMO", "Hûˋmoculture", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMOA", "Hûˋmoculture Flacon A", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMOB", "Hûˋmoculture Flacon B", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMOC", "Hûˋmoculture Flacon C", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMOD", "Hûˋmoculture Flacon D", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="LIQBIO", "Liquide Biologique", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="MUCO", "Mucoviscidose", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="HEMOA", "Hûˋmoculture Flacon A", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="ORLC", "Oto-rhino-laryngologie", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="OSTEOC", "Infection osteo-articulaire", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="SELLE", "Selle", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="SUPE", "Suppuration prûˋlevûˋe par ûˋcouvillon", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="SUPP", "Suppuration profonde", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(str_to_upper(type_prlvt)=="BMR-BHR", "Dûˋpistage rectal", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="URINE", "Urine", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="OPH", "Ophtalmologie", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="GENIT", "Gûˋnital", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="LCR", "Liquide cûˋphalo-rachidien", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="RECHSP", "Prûˋlû´vement environnemental", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="RECHSPE", "Prûˋlû´vement environnemental", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="Surv selles nûˋonat", "Surveillance selles nûˋonatologie", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="catheter_arteriel", "Cathûˋter artûˋriel", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="BIOPSIE", "Biopsie", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="LNOBIB", "Liquide non biologique sur flacon B", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="LIQBIA", "Liquide biologique sur flacon A", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="LIQBIB", "Liquide biologique sur flacon B", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="PNATAL", "Pûˋrinatalogie", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="MAT", "Matûˋriel", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="OSTEO", "Infection ostûˋo-articulaire", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="OPHA", "Ophtalmologie sur flacon A", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="OPHB", "Ophtalmologie sur flacon B", type_prlvt)) %>%mutate(type_prlvt=ifelse(type_prlvt=="LNOBIA", "Liquide non biologique sur flacon A", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(type_prlvt=="VAS", "Vasculaire", type_prlvt)) %>%
  mutate(type_prlvt=ifelse(is.na(type_prlvt), "", type_prlvt)) 


if(!(params$env_samples=="")){
  tab <- tab %>%
    mutate(ipp=ifelse(ipp==params$env_samples, "Prûˋlû´vement d'environnement", ipp))
}
  

# Outputting table 
kable(tab, col.names = c("GLIMS", "Identifiant Patient (IPP)", "Date de prûˋlû´vement", 
                         "Type de prûˋlû´vement", "Espû´ce MALDI-ToF", "MLST Sequence Type",
                         "Cluster"),
      align = "llllccc", "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("striped", "scale_down", "HOLD_position", "repeat_header")) %>%
  column_spec(4, width = "15em")

rm(tab)
```

# Conclusion du clustering

```{r conclusion, results="asis"}
cat(readLines('conclusion.txt'), sep = '\n')
```

\newpage

# Informations complûˋmentaires

```{r only_2_samples_clone, include=TRUE, results='asis'}
# First thing to do, only keep clone with nb of samples > 2 
singletons <- names(which(table(reports[[ clust_dist ]]) == 1))
reports[[clust_dist]][which(reports[[clust_dist]] %in% singletons)] <- "singletons"

# Compute proportion of non-singleton/singletons
prop_singleton <- round(length(reports$genome[which(reports[[clust_dist]] == "singletons")]) * 100 / length(reports$genome), digits = 2) 
prop_non_singleton <- 100 - prop_singleton

# Compute number of SNPs between samples if it's two non-clonal samples
if((length(reports$genome) == 2) & (prop_non_singleton == 0)){
  nb_variants <- unique(as.numeric(reports$nb_snps))
  cat("Les deux souches analysûˋes ne sont pas clonales, elles sont distantes de ", nb_variants, " SNPs.")
}

rm(singletons, nb_variants)
```

La proportion de singletons est de `r prop_singleton` %, contre `r prop_non_singleton` % d'isolats clonaux.

N.B: Un singleton correspond û  un isolat qui n'appartient û  aucun cluster, il n'est clonal avec aucun autre isolat.

## Temporalitûˋ des clones

```{r prlvt, out.height="300pt"}
# Computing limits of histogram (earliest and latest months)
first_month <- floor_date(min(reports$sampling_date), "month") %m-% months(1)
last_month <- ceiling_date(max(reports$sampling_date), "month")
months_period <- interval(ymd(first_month),ymd(last_month)) %/% months(1)
if(months_period < 6){
  first_month <- as.Date("2025-06-01")
  last_month <- as.Date("2026-06-01")
}

# Break data by month, e.g one bin one month 
out <- reports %>%
  select(genome, as.name(clust_dist), sampling_date) %>%
  mutate(month = floor_date(sampling_date, "month"))

# Color palette
cbPalette <- c("#D2C9CA", brewer.pal(length(unique(reports[[ clust_dist ]]))-1, "RdYlBu"))
# Plotting as a barplot of month categories
if("singletons" %in% out[[clust_dist]]){
  g <- ggplot(data=out, aes(x=month, fill = relevel(as.factor(eval(as.name(clust_dist))), "singletons"))) +
  geom_bar(colour = "black", aes(text = paste0("Nombre: ", after_stat(count))), width = 20) +
  ggtitle("Histogramme de dates de prûˋlû´vement") +
  labs(x = "Date de prûˋlû´vement par mois", y = "Nombre d'ûˋchantillons", fill = "Cluster") +
  scale_x_date(date_breaks = "month", date_labels = "%m/%Y", date_minor_breaks = "month", 
               limits = c(first_month, last_month)) +
  scale_y_continuous(n.breaks = 10) +
  scale_fill_manual(values=cbPalette) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1, size=5))
} else {
  g <- ggplot(data=out, aes(x=month, fill = as.factor(eval(as.name(clust_dist))))) +
  geom_bar(colour = "black", aes(text = paste0("Nombre: ", after_stat(count))), width = 20) +
  ggtitle("Histogramme de dates de prûˋlû´vement") +
  labs(x = "Date de prûˋlû´vement par mois", y = "Nombre d'ûˋchantillons", fill = "Cluster") +
  scale_x_date(date_breaks = "month", date_labels = "%m/%Y", date_minor_breaks = "month", 
               limits = c(first_month, last_month)) +
  scale_y_continuous(n.breaks = 10) +
  scale_fill_manual(values=cbPalette[-1]) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1, size=7))
}

g

rm(first_month, last_month, out, g, earliest, latest, nb_samples, prop_singleton, prop_non_singleton, cbPalette, months_period)
```


## Distance gûˋnûˋtique ûˋvaluûˋe

\noindent

```{r dist_snps, results='hide', fig.env='none', fig.align='default'}
# Plotting SNP matrices for each minicluster analyzed in whole cluster
library(gridExtra)
library(ComplexHeatmap)
library(circlize)
library(grid)

path_to_results <- paste0("~/Projets/2.Coding_projects/3.R/1.Automated_reports/phylogeny/",
                          "cluster_", params$cluster_id, "/",
                          params$species)
# Loop on miniclusters
all_dir_in_path <- list.dirs(path = path_to_results, full.names = FALSE, recursive = FALSE)
minicl_dir <- all_dir_in_path[grep("minicluster_[0-9]", all_dir_in_path)]
# Function
plot_snp_matrix <- function(path_to_res, minicl, all_samples = FALSE){
  # Input data = matrix of SNPs
  if(all_samples==FALSE){
    path_to_snps <- paste0(path_to_res, "/",
                           minicl, "/",
                           "gubbins/core_snp_matrix.tsv")
    if(!(file.exists(path_to_snps))){
      path_to_snps <- paste0(path_to_res, "/",
                             minicl, "/",
                             "snippy/snp_matrix.tsv")
    } else { print("hey it works woohoo")}
  } else {
    path_to_snps <- paste0(path_to_res, 
                           "/phylogenetic_tree/snippy/snp_matrix.tsv")
  }
  
  # Reading data
  snps_df <- data.table(read.table(path_to_snps, sep = "\t", header = TRUE, row.names = 1))
  # Remove reference if any
  if("Reference" %in% colnames(snps_df)){
    snps_df <- snps_df %>% select(-Reference) # Removing Reference Colum
    snps_df <- snps_df %>% filter(!row_number() == (length(snps_df)+1)) # Removing Reference Line
  }
  # Missing tip
  # missing_sample <- which(colnames(snps_df)=="Epi.2590")
  # snps_df <- snps_df[, (missing_sample) := NULL]
  # snps_df <- snps_df[-missing_sample]
  # Change rownames into GLIMS
  colnames(snps_df) <- gsub("\\.", "-", colnames(snps_df))
  # here is the problem ! please do the analysis step by step!
  print(colnames(snps_df))
  tmp <- reports[genome %in% colnames(snps_df)]
  tmp <- tmp[,c("genome","glims")]
  # I wanna be sure how setnames is working
  setnames(snps_df, old = tmp$genome, new = tmp$glims)
  rm(tmp)
  rownames(snps_df) <- colnames(snps_df)
  # Convert dataframe as matrix
  snps_matrix <- as.matrix(snps_df)
  colnames(snps_matrix) <- gsub("\\.", "-", colnames(snps_matrix))
  rownames(snps_matrix) <- colnames(snps_matrix)
  # Pheatmap lower tri
  snps_matrix[upper.tri(snps_matrix)] <- NA
  snps_matrix_display <- snps_matrix
  snps_matrix_display[is.na(snps_matrix_display)] <- ""
  diag(snps_matrix_display) <- ""
  plot_title <- paste0("Matrice de SNPs du ", str_to_title(gsub("\\_", " ", minicl)))
  # Color function equivalent to breaks: [-1, 0, 15, max+1]
  col_fun <- colorRamp2(
    c(0, 15, 16, 1e6),
    c("red", "red", "#008000", "#008000")
  )
  
  # Create Heatmap
  ht <- Heatmap(snps_matrix,
                name = "SNPs",
                col = col_fun,
                cluster_rows = FALSE,
                cluster_columns = FALSE,
                rect_gp = gpar(col = NA),  # remove cell borders
                column_names_rot = 45,
                row_names_side = "left",
                na_col = "white",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  val <- snps_matrix[i, j]
                  
                  # Grey background for diagonal
                  if (i == j) {
                    grid.rect(x = x, y = y, width = width, height = height,
                              gp = gpar(fill = "#353839", col = NA))  # grey fill, no border
                  } else {
                    # Let ComplexHeatmap fill color normally for non-diagonal cells
                    grid.rect(x = x, y = y, width = width, height = height,
                              gp = gpar(fill = fill, col = NA))
                    
                    if (!is.na(val)) {
                      grid.text(label = val, x = x, y = y, gp = gpar(col = "black", fontsize = 8))#8
                    }
                  }
                },
                width = unit(10, "cm"),    
                height = unit(5, "cm"),
                row_names_gp = gpar(fontsize = 12), #12
                column_names_gp = gpar(fontsize = 12), #12
                heatmap_legend_param = list(
                  at = c(0, 15, 20),  # midpoints of green and red zones
                  labels = c("0", "15", "16+"),
                  title = "Distance \nen SNPs",
                  title_gp = gpar(fontsize = 14),   # title style
                  labels_gp = gpar(fontsize = 12)    
                )
  )
  # Draw with main title and custom size
  draw(ht,
       heatmap_legend_side = "right",
       column_title = plot_title,
       column_title_gp = gpar(fontsize = 15),
       padding = unit(c(2, 2, 2, 2), "mm"))
  #return(ht)
}

# Let's use the function
list_of_plots <- sapply(minicl_dir, plot_snp_matrix, path_to_res = path_to_results)
rm(ht, list_of_plots, gtables, cl_samples, all_dir_in_path, minicl_dir, path_to_results, plot_snp_matrix)
```

\newpage

## Classes d'AMR par cluster

Les gû´nes de rûˋsistance indiquûˋs sont uniquement les gû´nes localisûˋs sur le chromosome. Les gû´nes de rûˋsistance sur plasmide ne sont pas inclus dans ce rapport.

```{r amr_barplot_per_classes, include=FALSE}
# Creating a list of AMR class found per clone
amr_list <- reports %>%
  select(as.name(clust_dist), amr_classes) %>%
  mutate(amr_classes = tolower(amr_classes))
amr_list <- separate_longer_delim(amr_list, amr_classes, ",")

# Histogram of AMR_classes (aminoglycoside, phenicol, etc.)
g <- ggplot(data=amr_list, aes(x=amr_classes)) +
  geom_bar(stat="count", fill = "lightblue", aes(text = paste0("Count: ", after_stat(count)))) +
  ggtitle("AMR classes") +
  labs(x = "AMR class", y = "Count") +
  scale_y_continuous(breaks= pretty_breaks()) +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust = 1))

g

rm(g, amr_list)
```

```{r new_amr}
# Loading Data with AMRFinder report [GLOBALLY]
load(params$path_to_scan)
rm(amrfinder_reports, contig_reports, mge_reports, mlst_reports, quast_reports, sourmash_reports)
```

```{r amr_class_clone, results='asis', message=FALSE, warning=FALSE}
# Display table of ARGs on chromosome. 
library(dplyr)
library(stringr)
library(knitr)
library(kableExtra)
library(purrr)

rows_per_table <- 30  # <- adjust page size

clones <- sort(unique(reports[[clust_dist]]))

for (c in clones) {
  if (identical(c, "singletons")) {
    cat("### Singletons\n\n")
    cat("Au sein des **singletons**, on retrouve ces gû´nes de rûˋsistances aux antibiotiques :\n\n")
  } else {
    cat("### Cluster ", c, "\n\n", sep = "")
    cat("Au sein du **cluster ", c, ",** on retrouve ces gû´nes de rûˋsistances aux antibiotiques :\n\n", sep = "")
  }

  # Which genomes are in my clusters ?
  samples <- reports$genome[reports[[clust_dist]] == c]

  # Get AMR report of those genomes (data.table subset)
  amr_report <- combined_amr_report[`Element type` == "AMR" & genome %in% samples]

  # Loading only replicon AMR genes
  list_arg <- amr_report[molecule_type == str_to_lower(params$replicon)]

  # Arranging AMR results
  out <- list_arg %>%
    as.data.frame() %>%  # keep dplyr happy even if input is data.table
    select(genome, `Gene symbol`, `Sequence name`, Class, Subclass) %>%
    distinct() %>%
    group_by(`Gene symbol`, `Sequence name`, Class, Subclass) %>%
    mutate(Class = str_to_title(Class), Subclass = str_to_title(Subclass)) %>%
    summarise(Pourcentage = round(n() * 100 / length(samples), 2), .groups = "drop") %>%
    arrange(desc(Pourcentage)) %>%
    mutate(Pourcentage = paste0(Pourcentage, " %"))

  # Skip if nothing to print for this cluster
  if (nrow(out) == 0) {
    cat("_Aucun gû´ne AMR dûˋtectûˋ pour ce groupe._\n\n")
  } else {
    # Display on several pages if needed
    # Split into chunks of rows_per_table and print each as a kable
    chunks <- split(out, ceiling(seq_len(nrow(out)) / rows_per_table))
    imap(chunks, ~{
      if (length(chunks) > 1) cat(paste0("**Partie ", .y, " / ", length(chunks), "**\n\n"))
      kable(.x,
            format = "latex", booktabs = TRUE,
            col.names = c("Nom du gû´ne", "Description", "Classe d'AMR", "Sous-classe d'AMR", "Pourcentage de prûˋsence"),
            align = "llllr",
            escape = TRUE  # set FALSE only if you need LaTeX in cells
      ) %>%
        kable_styling(latex_options = c("striped", "scale_down", "HOLD_position")) %>%
        column_spec(2, width = "20em") %>%
        column_spec(5, width = "8em") %>%
        print()
      cat("\n\n")
      cat("\\newpage\n\n")
    })
  }

  # Page break BETWEEN clusters (PDF only)
  if (knitr::is_latex_output()) cat("\\newpage\n\n")
}

# clean
rm(amr_report, list_arg, out, c, clones, samples)
```

```{r tree, echo=FALSE, results='asis'}
# out.height="300pt"
# samples_clustered <- reports[[clust_dist]][which(reports[[clust_dist]] != "singletons")]
# if(length(reports$genome) > 2){
#   cat("\\newpage \n\n")
#   cat("## Phylogûˋnie \n\n")
#   cat("A partir de l'alignement des SNPs dûˋtectûˋs dans les ûˋchantillons, un arbre phylogûˋnûˋtique du cluster est infûˋrûˋ.\n\n")
#   if(length(samples_clustered)>0){
#     cat("Les isolats clonaux sont coloriûˋs d'une couleur spûˋcifique.\n\n")
#   }
#   path_to_tree <- paste0("../phylogeny/cluster_",params$cluster_id,"/",params$species,
#                          "/phylogenetic_tree/clonalframeml/GTR_tree.jpeg")
#   i <- include_graphics(path_to_tree)
#   cat("\\begin{center}")
#   cat("\\includegraphics[width=5in,height=7.5in]{", path_to_tree, "}")
#   cat("\\end{center}")
#   rm(path_to_tree)
# }
# rm(out.height)
```

```{r cleaning}
rm(global_theme, params, reports, species, nb_snps, clust_dist, samples_clustered, chunks, combined_amr_report, mobtyper_reports, clust_levels, ggplot_title, nb_patients, rows_per_table)
```






